---
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
---




```{r setKnitrOptions, include=FALSE}
knitr::opts_chunk$set(results='asis', echo = FALSE, error = TRUE, cache = FALSE, cache.comments = TRUE, fig.dim = c(7,7), autodep = TRUE)
```




```{r loadParamsAndDependencies, warning=FALSE, message=FALSE}

### Load general Deps
library( RColorBrewer)

cat("<BR>WORKING_DIR=", WORKING_DIR)
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepare_data.R"));
read_chunk( path=file.path( WORKING_DIR, "02_compare_markers.R"));
read_chunk( path=file.path( WORKING_DIR, "03_find_tissue_markers.R"));
read_chunk( path=file.path( WORKING_DIR, "04_prepare_paper_figures.R"));
#read_chunk( path=file.path( WORKING_DIR, "90_finalize.R"));
```

---
title: `r paste(SAMPLE_NAME, PROJECT_NAME, LITERAL_TITLE, sep ="<BR>")`
---


```{r initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( paramsEnv, algo="murmur32");

# Create a 'safe' unique prefix for output files
outputFilesPrefix = path_sanitize( gsub("([[:punct:]])|\\s+", 
                                        "_", 
                                        paste0(PROJECT_NAME, "_",
                                               SAMPLE_NAME, "_",
                                               startTimeFileName, "_",
                                               paramsHash, "_")),
                                   replacement = "_");

# Set seed for pseudo random number generator
set.seed(42);

# Create a hidden empty heatmap to prevent heatmapr interference with plotly functions (bug ?)
cat('<div style="display: none">');
main_heatmap(matrix(0));
cat('</div>');
```




## Compare tissues


### Run informations {.tabset .tabset-fade}

#### Run
<center>
```{r runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Sample'   = SAMPLE_NAME,
           'Hash'     = paramsHash));
```
</center>

#### Parameters
```{r runInfo_params}
#showSimpleDT( paramsEnv); # Does not handle lists or nested lists easily
listParams = as.list( paramsEnv);
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

#### System
<center>
```{r runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

#### R
```{r runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
### {.toc-ignore} 
<br /><br />


### Read data from Heterogenity by Tissue analysis
```{r dataLoad, out.width='50%'}
print( datatable( NULL))
<<load_data>>
```
<br/><br />

### Compare clusters between tissues

In this section we look at the marker genes found between cluster of different tissues. For each pair of tissues we look at each pair of clusters (one
per tissue). For each cluster pair, we compute the Jaccard Index :
<br><br>
Jaccard Index = #(intersection of marker genes of the two clusters) / #(union of the marker genes of the two clusters).
<br><br>
The result is presented as a heatmap. The color is relative to the Jaccard Index. The text represent (i) the number of marker genes in the intersection and (ii)
the percentage of the intersection compare to the union (i.e. the Jaccard index multiplied by 100).<br>
The next step search, for each cluster of one tissue, its best match (best Jaccard Index) through the clusters of the other tissue. The list of
marker genes at the intersection of the list of marker genes of each cluster is then shown. Note that the list of gene is order by increasing adjusted p-value.

```{r CompareClusters, out.width='50%', out.height='50%'}
<<compare_markers_between_clusters>>
```

### Identify marker genes between tissues

In this section, we consider each the tissue as a whole. We remove cells that were identified as affected by dissociation process and remove BCR genes. Then
we look at markers genes for each tissue.

```{r identifyTissueMarkerGenes, out.width='50%', out.height='50%'}
<<identify_tissue_marker_genes>>
```

## Prepare paper figures
```{r preparePaperFigureHeatmap}
<<prepare_paper_figure_heatmap>>
```
<br/><br />


