---
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
---




```{r setKnitrOptions, include=FALSE}
knitr::opts_chunk$set(results='asis', echo = FALSE, error = TRUE, cache = FALSE, cache.comments = TRUE, autodep = TRUE)
```




```{r loadParamsAndDependencies, warning=FALSE, message=FALSE}
### Create output dir
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE, recursive = TRUE)

### Load general Deps
cat("<BR>WORKING_DIR=", WORKING_DIR)
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"));
read_chunk( path=file.path( WORKING_DIR, "02_mergeData.R"));
read_chunk( path=file.path( WORKING_DIR, "03_analyzeClonotypes.R"));
read_chunk( path=file.path( WORKING_DIR, "04_analyzePhenotypes.R"));
read_chunk( path=file.path( WORKING_DIR, "05_analyzeAGspecificity.R"));
read_chunk( path=file.path( WORKING_DIR, "06_analyzeExpression.R"));
read_chunk( path=file.path( WORKING_DIR, "07_analyzeSpecificClonotypes.R"));
#read_chunk( path=file.path( WORKING_DIR, "90_finalize.R"));
```

---
title: `r paste(PROJECT_NAME, LITERAL_TITLE, sep ="<BR>")`
---


```{r initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( paramsEnv, algo="murmur32");

# Create a 'safe' unique prefix for output files
outputFilesPrefix = path_sanitize( gsub("([[:punct:]])|\\s+", 
                                        "_", 
                                        paste0(PROJECT_NAME, "_",
                                               startTimeFileName, "_",
                                               paramsHash, "_")),
                                   replacement = "_");

# Set seed for pseudo random number generator
set.seed(42);

# Create a hidden empty heatmap to prevent heatmapr interference with plotly functions (bug ?)
cat('<div style="display: none">');
main_heatmap(matrix(0));
cat('</div>');
```




## BCR Analysis


### Run informations {.tabset .tabset-fade}

#### Run
<center>
```{r runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Hash'     = paramsHash));
```
</center>

#### Parameters
```{r runInfo_params}
#showSimpleDT( paramsEnv); # Does not handle lists or nested lists easily
listParams = as.list( paramsEnv);
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

#### System
<center>
```{r runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

#### R
```{r runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
### {.toc-ignore} 
<br /><br />


### Load and filter data produced by BCR reconstruction and analysis
```{r prepareData, results="asis", fig.keep="all", echo=FALSE, message=FALSE, warning=TRUE, fig.dim = c(7,7)}
<<prepare_bcr_data>>
```
<br /><br />


### BCR repertoire dispersion analysis on all tissues

Here we look at the distribution of clonotype groups on the complete data. Here are the displayed analysis:

* the clusters on a UMAP embedding for reference
* the dispersion of clonotypes across the samples : as a table and as a cumulative barplot of percentages
* the dispersion of the samples across the 11 largest clonotypes: as a table and as a cumulative barplot of percentages
* the dispersion of the clusters across the 11 largest clonotypes: as a table and as a cumulative barplot of percentages
* the dispersion of the clusters across the isotype of heayvy-chain (all cells): as a table and as a cumulative barplot of percentages
* the dispersion of the clusters across the isotype of light-chain (all cells): as a table and as a cumulative barplot of percentages
* the dispersion of the clusters across the V-Segment sub-group of heayvy-chain (all cells): as a table and as a cumulative barplot of percentages
* the dispersion of the clusters across the V-Segment sub-group of light-chain (all cells): as a table and as a cumulative barplot of percentages
* the dispersion of the clusters across the CDR3 lengths of heayvy-chain (all cells): as a table and as a cumulative barplot of percentages
* the dispersion of the clusters across the CDR3 lengths of light-chain (all cells): as a table and as a cumulative barplot of percentages
* the dispersion of the clusters across the number of mutations of heayvy-chain (all cells): as a table and as a cumulative barplot of percentages
* the dispersion of the clusters across the number of mutations of light-chain (all cells): as a table and as a cumulative barplot of percentages

<b>Important note : barplots of cumulative percentages have to be crossed with the corresponding tables to consider the percentages with the numbers they are based on.</b>

```{r analyzeClonotypesAllTissues, results="asis", fig.keep="all", echo=FALSE, message=FALSE, warning=TRUE, fig.dim = c(7,7)}
custom.rna.seurat = readRDS( file.path( PATH_SEURAT_FILTERED_OBJECT, paste0( PROJECT_SHORT_NAME, "_custom.rna.seurat.RDS")))
<<merge_data>>
<<analyze_clonotypes>>
```

### Analysis of the relation between BCR repertoire and Index sort phenotypes in two dimensions

```{r analyzeBCRRepertoireToPhenotype, results="asis", fig.keep="all", echo=FALSE, message=FALSE, warning=TRUE}
<<analyze_bcrrepertoire_to_phenotype>>
```


### Analysis of the antigen specificity

```{r analyzeAGspecificity, results="asis", fig.keep="all", echo=FALSE, message=FALSE, warning=TRUE}
<<analyze_AGspecificity>>
```

### Analysis of the expression of pehnotype groups

```{r analyzePhenotypeToExpression, results="asis", fig.keep="all", echo=FALSE, message=FALSE, warning=TRUE}
<<analyze_phenotype_to_expression>>
```

### Analysis of antigen specificy of the cells using the top N used V-genes

In this section, we select the top N used heavy-chain V-genes and the top N used ligh-chain V-genes and look at the Ag Specificity, the mouse of origin, the clonotype size and the cluster of the cells using those gene.

For each indicator, the following tables show the number of cells per AG affinity group in function of the number N of top v-genes. The first table is the raw counts, the second the percentage per line and the third the percentage in the Ag-specificity groups.

Each table is shown in a heatmap. The last heatmap showw the Pearson residual of the chi2 test on the first table.


```{r analyzeSpecificClonotypes, results="asis", fig.keep="all", echo=FALSE, message=FALSE, warning=TRUE}
<<analyze_specific_clonotypes>>
```

