---
output:
  html_document: 
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
---




```{r setKnitrOptions, include=FALSE}
knitr::opts_chunk$set( results='asis', echo = FALSE, error = TRUE, cache = FALSE, cache.comments = TRUE, fig.dim = c(7,7), autodep = TRUE)
```




```{r loadParamsAndDependencies, warning=FALSE, message=FALSE}
# Create the analysis output dir 
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE, recursive = TRUE)

### Load general Deps
cat("<BR>WORKING_DIR=", WORKING_DIR)
source( file.path( WORKING_DIR, "00_generalDeps.R"));

# Import chunks from source files
read_chunk( path=file.path( WORKING_DIR, "01_prepareData.R"));
read_chunk( path=file.path( WORKING_DIR, "02_variableGenes.R"));
read_chunk( path=file.path( WORKING_DIR, "03_cellHeterogeneity.R"));
```

---
title: `r paste(SAMPLE_NAME, PROJECT_NAME, "Heterogeneity analysis through all samples", sep =" - ")`
---


```{r initFunctions}

### Create output directory
dir.create( PATH_ANALYSIS_OUTPUT, showWarnings = FALSE);


### Record starting time (for report and file names)
startTime = Sys.time();
startTimeReport   = format( startTime, "%Y %m %d %H:%M:%S")
startTimeFileName = format( startTime, "%Y_%m_%d_%H_%M_%S")


### Initialize utils

# Create a hash for all parameters
paramsHash = digest( paramsEnv, algo="murmur32");

# Create a 'safe' unique prefix for output files
outputFilesPrefix = path_sanitize( gsub("([[:punct:]])|\\s+", 
                                        "_", 
                                        paste0(PROJECT_NAME, "_",
                                               SAMPLE_NAME, "_",
                                               startTimeFileName, "_",
                                               paramsHash, "_")),
                                   replacement = "_");

# Set seed for pseudo random number generator
set.seed(42);

# Create a hidden empty heatmap to prevent heatmapr interference with plotly functions (bug ?)
cat('<div style="display: none">');
main_heatmap(matrix(0));
cat('</div>');
```




## Heterogeneity analysis on all samples


### Run informations {.tabset .tabset-fade}

#### Run
<center>
```{r runInfo_run}
pander( c( 'Date'     = startTimeReport,
           'Sample'   = SAMPLE_NAME,
           'Hash'     = paramsHash));
```
</center>

#### Parameters
```{r runInfo_params}
#showSimpleDT( paramsEnv); # Does not handle lists or nested lists easily
listParams = as.list( paramsEnv);
pander(listParams[order(names(listParams))]); # Show params in alphabetical order
```

#### System
<center>
```{r runInfo_sysInfo}
showSimpleDT( Sys.info())
```
</center>

#### R
```{r runInfo_r}
pander( sessionInfo());
```

<!-- Ending the tabset section -->
### {.toc-ignore} 
<br /><br />




### Load data from QC/Demultiplex analysis
```{r dataLoad}
<<loadData>>
```
<br /><br />

### Variable genes
```{r findVariableGenes}
<<findVariableGenes_seuratMethod>>
```

Number of variable genes found: `r variablesGenesStats`

```{r findVariableGenes_summaryPlot}
<<findVariableGenes_summaryPlot>>
```
<br />

Top `r VARIABLE_FEATURES_SHOWTOP` most variable annotations:

```{r findVariableGenes_summaryTable}
<<findVariableGenes_summaryTable>>
```
<br /><br />



### Clusters identification

#### Cells distribution
```{r heterogeneity_identifyClusters}
<<heterogeneity_identifyClusters>>
```

#### Statistics by cluster
```{r heterogeneity_identifyClusters_splitStats}
<<heterogeneity_identifyClusters_splitStats>>
```
<br /><br />


### PCA results

PCA projections for dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca, fig.dim = c(3.5,3.5), out.width='33%'}
<<heterogeneity_pca>>
```
<br />

UMIs counts on PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca_umisCounts, fig.dim = c(3.5,3.5), out.width='33%'}
<<heterogeneity_pca_umisCounts>>
```
<br />

Genes counts on PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca_genesCounts, fig.dim = c(3.5,3.5), out.width='33%'}
<<heterogeneity_pca_genesCounts>>
```
<br />

Correlation (Spearman) of PCA dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")` with UMIs and genes counts  
```{r heterogeneity_pca_correlations, fig.dim = c(9,6)}
<<heterogeneity_pca_correlations>>
```
<br />

PCA loadings of top `r PCA_PLOTS_NBFEATURES` features for dimensions `r paste( 1:PCA_PLOTS_NBDIMS, collapse="/")`  
```{r heterogeneity_pca_loadings, fig.dim = c(3.5,3.5), out.width='33%'}
<<heterogeneity_pca_loadings>>
```
<br /><br />




### Dimensionality reduction and clustering {.tabset .tabset-fade}
```{r heterogeneity_dimReduc}
<<heterogeneity_dimReduc>>
```

#### tSNE {.tabset .tabset-fade}
```{r heterogeneity_dimReduc_plot_tSNE}
useReduction='tsne';

<<heterogeneity_dimReduc_interactivePlot_clusters>>
#<<heterogeneity_dimReduc_interactivePlot_samples>>
#<<heterogeneity_dimReduc_details_samples>>
rm("useReduction");
```

#### UMAP {.tabset .tabset-fade}
```{r heterogeneity_dimReduc_plot_UMAP}
useReduction='umap'

<<heterogeneity_dimReduc_interactivePlot_clusters>>
#<<heterogeneity_dimReduc_interactivePlot_samples>>
#<<heterogeneity_dimReduc_details_samples>>
rm("useReduction");
```


<!-- Ending the tabset section -->
#### {.toc-ignore}

<!-- Ending the tabset section -->
### {.toc-ignore}
<br /><br />




### Marker genes
```{r heterogeneity_markerGenes}
<<heterogeneity_markerGenes>>
```

#### Marker genes list
```{r heterogeneity_markerGenes_table}
<<heterogeneity_markerGenes_table>>
```
<br />

#### Marker genes heatmap

Expression heatmap of top marker genes sorted by Seurat clusters:

```{r heterogeneity_markerGenes_heatmap}
<<heterogeneity_markerGenes_heatmap>>
```

#### Marker genes expression on t-SNE/UMAP {.tabset .tabset-fade}

##### tSNE {.tabset .tabset-fade}
```{r heterogeneity_markerGenes_expression_tSNE, fig.dim = c(3,3), out.width='25%'}
useReduction='tsne';

cat("\n<a href='#dimensionality-reduction'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_markerGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}

##### UMAP {.tabset .tabset-fade}
```{r heterogeneity_markerGenes_expression_UMAP, fig.dim = c(3,3), out.width='25%'}
useReduction='umap';

cat("\n<a href='#dimensionality-reduction'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_markerGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}


<!-- Ending the tabset section -->
#### {.toc-ignore}
<br /><br />


#### Marker genes expression on violin plots

Here we show the detail of expression of all genes identified as marker for at least one cluster
as violin plot per cluster (with jitter point). Numbers above each violin plot is (i) top: the number
of cells with a positive expression of the gene in the cluster (ii) bottom : the number of cells
with a null expression of the gene in the cluster.

```{r heterogeneity_markerGenes_expression_violin, fig.keep = "all", fig.dim = c(3,3), out.width='25%'}
<<heterogeneity_markerGenes_expression_violinplot>>
```


### Monitored genes
```{r heterogeneity_monitoredGenes}
<<heterogeneity_monitoredGenes>>
```

#### Monitored genes heatmap

Expression heatmap of monitored genes sorted by Seurat clusters:

```{r heterogeneity_monitoredGenes_heatmap}
<<heterogeneity_monitoredGenes_heatmap>>
```

#### Monitored genes expression {.tabset .tabset-fade}

##### tSNE {.tabset .tabset-fade}
```{r heterogeneity_monitoredGenes_expression_tSNE, fig.dim = c(3,3), out.width='25%'}
useReduction='tsne'

cat("\n<a href='#dimensionality-reduction'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_monitoredGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}

##### UMAP {.tabset .tabset-fade}
```{r heterogeneity_monitoredGenes_expression_UMAP, fig.dim = c(3,3), out.width='25%'}
useReduction='umap'

cat("\n<a href='#dimensionality-reduction'>");
<<heterogeneity_dimReduc_thumbnail>>
cat("</a> \n \n");

<<heterogeneity_monitoredGenes_expression>>

rm("useReduction");
```

<!-- Ending the tabset section -->
##### {.toc-ignore}

<!-- Ending the tabset section -->
#### {.toc-ignore}
<br /><br />

#### Monitored genes expression on violin plots

Here we show the detail of expression of all monitored genes
as violin plot per cluster (with jitter point). Numbers above each violin plot is (i) top: the number
of cells with a positive expression of the gene in the cluster (ii) bottom : the number of cells
with a null expression of the gene in the cluster.

```{r heterogeneity_monitoredGenes_expression_violin, fig.keep = "all", fig.dim = c(3,3), out.width='25%'}
<<heterogeneity_monitoredGenes_expression_violinplot>>
```


### Analysis of expresison biais on monitored and marker genes

In this section we have a look to the distribution of gene expression mean versus percentage of zero counts.

```{r heterogeneity_monitoredAndMarkerGenes_stats, fig.keep = "all", out.width='100%'}
heterogeneity_monitoredAndMarkerGenes_stats
```


### Behavior of negative eYFP cells

In this section we pay attention to some stats on the eYFP genes to try to identify some specific behavior on eYFP negative cells.

```{r heterogeneity_eYFPGene_stats, fig.keep = "all", fig.dim = c(2,2), out.width='50%'}
heterogeneity_eYFPGene_stats
```


<!-- Ending the tabset section -->
#### {.toc-ignore}

<!-- Ending the tabset section -->
### {.toc-ignore}
<br /><br />




```{r final_saveSessionImage}
#<<final_saveSessionImage>>
```
